Automatically Exploring Hypothesis Spaces

### 2018-03-18 topics.tex
prettier rules

\usepackage{amsmath}
\usepackage[margin=0.1in]{geometry}
\usepackage{booktabs}
\usepackage{tipa}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{tikz}
\usetikzlibrary{fit}
%\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{amsfonts}       % blackboard math symbols

 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

@@

The goal of the grammar learner, whether it is a child, linguist, or computer program, is to
infer the grammar that gave rise to
a collection of utterances.
One way of
modeling this inference problem is to (1) place a prior distribution over grammars, for example, a distribution that puts more probability mass on shorter or simpler grammars; (2) equip each grammar with a
\emph{likelihood model} that specifies exactly how likely a grammar is to produce a given utterance;
and then (3) use Bayes's rule to work backwards from the
utterances to the grammar that was likely to produce them:
\begin{equation}
  \probability\left[\text{grammar}|\text{utterances} \right]\propto\underbrace{\probability\left[\text{utterances}|\text{grammar} \right]}_{\text{Likelihood}}\times\underbrace{\probability\left[\text{grammar} \right]}_{\text{Prior}}\label{posterior}
\end{equation}
We can think of the above equation as a recipe for
scoring how well a grammar explains a
collection of utterances.
Once we have a prior and a likelihood,
we know how to evaluate how well a grammar explains the data, modulo the assumptions implicit in the
prior and likelihood.

In this work we will use this probabilistic framing
to explore spaces of grammars
that an AGL learner could use to explain
the pronunciations of words.
Here the grammars will be SPE-style rewrite rules~\cite{chomsky1968sound},
commonly used in generative phonology,
but at a high level this framing can apply to
learning other aspects of grammar like semantics~\cite{piantadosi2011learning}, syntax~\cite{perfors2011learnability}, and morphology~\cite{tim}.

A simple and intuitive prior over grammars is one which penalizes longer grammars and favors parsimonious grammars, for example defining $\probability\left[\text{grammar} \right]\propto\exp\left(- \text{\# symbols in grammar} \right)$. An example of a likelihood model,
and the one we use here,
is just to count the number of extra bits or symbols needed to encode an utterance given the grammar\footnote{This is equivalent to a minimum description length (MDL) interpretation}.

@@

The fundamental hypothesis underlying AGL research is that
artificial grammar learning must engage some shared resource with first language acquisition --
and so we can gain insights into first language acquisition by
instead studying AGL in controlled experiments.
Our contribution here is the demonstration that
a single BPL model can explain how a learner can quickly infer a grammar represented as SPE--style rules.
We collected grammars from AGL datasets~\cite{gerken2010infants,marcus1999rule,frank2011three}
and presented our BPL learner with data drawn from these grammars (Table~\ref{artificialGrammarTable}),
finding that our model automatically discovers the grammars that
 underly these data sets.
\begin{table}[h!]
\centering\begin{tabular}{lcll}
  \toprule
  Grammar&Example input to learner&Grammar we infer&Natural language analogues\\\midrule
  \begin{tabular}{l}
    ABA (same/different/same)\\
  Marcus et al. 1999.     
    \end{tabular}&
  \begin{tabular}{l}
    wofewo\\
    lovilo\\
    fimufi
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / #_|$\sigma \sigma_i$
  \end{tabular}& Reduplication (eg, Tagalog)
  \\\midrule
  \begin{tabular}{l}
      ABB (different/same/same)\\
  Marcus et al. 1999. 
    \end{tabular}&
  \begin{tabular}{l}
        wofefe\\
    lovivi\\
    fimumu
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / |$\sigma_i$\verb|_#|
  \end{tabular}&
  Reduplication (eg, Tagalog)\\\midrule
  \begin{tabular}{l}
      ABx (different/different/constant)
    \end{tabular}&
  \begin{tabular}{l}
        wofeka\\
    lovika\\
    fimuka
  \end{tabular}&
  \begin{tabular}{l}
stem$ + x$
  \end{tabular}&concatenative morphology
  \\\midrule
  \begin{tabular}{l}
    AAx (same/same/constant)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wowoka\\
    loloka\\
    fifika
  \end{tabular}&
  \begin{tabular}{l}
    stem + $x$\\
$\varnothing\to \sigma_i$\verb| / #_|$\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    reduplication\\concatenative morphology
    \end{tabular}\\\midrule
  \begin{tabular}{l}
    AxA (same/constant/same)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wokawo\\
    lokalo\\
    fikafi
  \end{tabular}&
  \begin{tabular}{l}
$x + $ stem\\
    $\varnothing$ $\to\sigma_i$  / \verb|# _| $\sigma$ $\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    Infixing\\Reduplication
    \end{tabular}
  \\\midrule
  \begin{tabular}{l}
    Pig Latin
    \end{tabular}&
  \begin{tabular}{l}
    \textipa{pIg}$\to$\textipa{Igpe}\\
    \textipa{latIn}$\to$\textipa{atIle}\\
    \textipa{\ae sk}$\to$\textipa{\ae ske}\\
  \end{tabular}&
  \begin{tabular}{l}
      $\varnothing\to$\verb|C|$_i$ $/$ \verb|#C|$_i$\verb| [ ]* _ #|\\%
      $\varnothing\to$ \textipa{e} $/$ \verb| _ #|\\%
 \verb|C|$\to\varnothing$ $/$ \verb|# _|
  \end{tabular}  & \begin{tabular}{l}
    \textbf{Tim, is there an  analog?}\\
    \textbf{word initial segments moves to  }\\
    \textbf{word final?}
    \end{tabular}
  \bottomrule  \end{tabular}
\caption{Using BPL to model artificial grammar learning. Learner given 5 examples.}\label{artificialGrammarTable}
\end{table}

@@

As an example of how the Pareto front offers a lens on the
space of possible hypotheses,
consider the upper left Pareto front in figure~\ref{quadrupleFrontier} (\emph{aab, 1 example}).
Here the learner has seen a single word, [\textipa{vEfefe}].
Some grammars living on the Pareto frontier are:
\begin{itemize}
\item \textbf{A grammar that does nothing:} In the lower right-hand corner of~\ref{quadrupleFrontier}
  is the grammar labeled ``surface=underlying'',
  which just says that every word (a ``surface pronunciation'')
  is represented (``underlyingly'') literally how it is pronounced.
  The observed word [\textipa{vEfefe}] is represented as /\textipa{vEfefe}/, which has 6 symbols,
  giving a fit to the data of -6.
\item \textbf{A grammar that duplicates syllables:} Highlighted in dark red in~\ref{quadrupleFrontier}
  is a grammar with the single rule $\varnothing\to\sigma_i/\_\sigma_i$\texttt{\#},
  which inserts a copy of the last syllable syllable.
  This rule generates the word [\textipa{vEfefe}] by starting with the stem (i.e., underlying form) /\textipa{vEfe}/ (which has 4 symbols, fit to data of -4) and then applying the rule in the grammar,
  which copies the last syllable and makes [\textipa{vEfefe}].
\item \textbf{A grammar that duplicates syllables and appends
  morphemes:} Highlighted in pink in~\ref{quadrupleFrontier} are
  grammars that duplicate a syllable but also append or prepend extra
  morphemes.  These correspond to generalizations where the learner
  believes that different parts of [\textipa{vEfefe}] correspond to
  prefixes or suffixes in the language.  For example, the grammar in
  the upper left corner incorporates the suffix /\textipa{fe}/ as well
  as the prefix /\textipa{vE}/, and explains the observation using an
  underlying form that is completely empty (fit to data of 0) -- this
  grammar has memorized the observed data, and so maximally compresses
  it, but at the cost of having many symbols inside of the grammar (22 symbols, vs 6 symbols in the grammar that just duplicates a syllable).  
  \end{itemize}

@@

\includegraphics[width = 0.5\textwidth]{abb1.png} \includegraphics[width = 0.5\textwidth]{abb3.png}
\includegraphics[width = 0.5\textwidth]{aax1.png} \includegraphics[width = 0.5\textwidth]{aax3.png}

@@

These natural language results are still pulmonary and
ongoing. Figure~\ref{naturalLanguageExample} shows our model solving a
representative textbook problem.  We believe this extension two
natural language is an important next step for models of artificial
grammar learning, and helps deliver on the basic promises and
presuppositions of AGL research.

\begin{figure}[H]\centering
  \begin{minipage}{7cm}\centering
    \textbf{Textbook Tibetan counting problem:}
    \begin{tabular}{ll}\toprule
Number & Surface form
\\ \midrule
10 & \textipa{d\super Zu}\\
1 & \textipa{d\super Zig}\\
11 & \textipa{d\super Zugd\super Zig}\\
4 & \textipa{Si}\\
%14 & \textipa{d\super ZubSi}\\
40 & \textipa{Sibd\super Zu}\\
9 & \textipa{gu}\\
19 & \textipa{d\super Zurgu}\\
%90 & \textipa{gubd\super Zu}\\
5 & \textipa{Na}\\
%15 & \textipa{d\super ZuNa}\\
%50 & \textipa{Nabd\super Zu}\\
\bottomrule
\end{tabular}
  \end{minipage}%
  \hspace{0.5cm}%
  \begin{minipage}{6cm}\centering
    \textbf{Our model's solution:}

    Induced phonological rule: \texttt{C}$\to\varnothing /$\texttt{\#\_C} \\
    \hspace{0.5cm}\emph{(Reduce word initial consonant clusters)}

@@

%\pagebreak
%% \begin{figure}
%%   \begin{tikzpicture}
%%     %  \draw[help lines] (0,0) grid (10,-10);
%%     \node(UG)[purple,align = center,rotate = 90] at (1,-1) {\textbf{Universal}\\\textbf{Grammar}};
%%     %\draw[ultra thick] (2,0) rectangle (13,-2.3);

%%     \node(sounds) at (7,-0.3) {Phoneme Inventory};
%%     \node(soundExamples)[draw,align = center] at (7,-1.3) {\textipa{@, p, E, 2, k}\\\textipa{a, 3, A, O, d}\\$\cdots$};

%%     \node at (11,-0.3) {Program space};
%%     \node(programSpace)[draw,align = left] at (11,-1.3) {$\mathcal{M}\to \mathcal{A} | $\verb|stem|$|\mathcal{M}+\mathcal{M}$\\$\mathcal{P}\to \mathcal{R}\circ \mathcal{R}\circ \cdots \circ \mathcal{R}$};

%%     \node(featureSpace) at (3.5,-0.3) {Feature space};
%%     \node(featureExamples)[draw,align = center] at (3.5,-1.3) {\verb|voiced|, \verb|vowel|\\\verb|obstruent|, \verb|nasal|\\$\cdots$};

%%     \draw[thick,->] (featureExamples.east) -- (soundExamples.west);
%%     \draw[thick,->] (soundExamples.east) -- (programSpace.west);

%%     \node(universalBox)[purple,draw,ultra thick, fit = (sounds) (featureSpace) (featureExamples) (soundExamples) (programSpace)] {};
%%     \draw(lowerCausalArrow)[thick,->] (featureExamples.south) .. controls (7,-2.5) ..  (programSpace.south);

%%     \node(morphology)[blue,align = above,rotate = 90] at (1,-5) {\textbf{Programs}};
%%     \draw[ultra thick,->] (UG.west) -- (morphology.east);

%%     \node(ABAmorphologyLabel) at (10.5,-3.5) {Pig Latin Morphology};
%%     \node(ABAmorphology) [draw,align = left] at (10.5,-4){ \verb|observation|$\to$\verb|stem|};

%%     \node(ABAphonologyLabel) at (10.5,-5.3) {Pig Latin Phonology};
%%     \node(ABAphonology) [draw,align = left] at (10.5,-6.3) {%
%% $r_1$:      $\varnothing\to$\verb|C|$_i$ $/$ \verb|#C|$_i$\verb| [ ]* _ #|\\%
%%       $r_2$:      $\varnothing\to$ \textipa{e} $/$ \verb| _ #|\\%
%%     $r_3$: \verb|C|$\to\varnothing$ $/$ \verb|# _|};
%%     \node(ABAmorphologyBox)[blue,draw,ultra thick, fit = (ABAmorphologyLabel) (ABAphonology)] {};

%%     %  \draw[ultra thick] (2,-3) rectangle (7,-7);

%%     \node(polishMorphologyLabel) at (4.5,-3.5) {Polish Morphology};
%%     \node(polishMorphology) [draw,align = left] at (4.5,-4.3) {\verb|singular|$\to$ \verb|stem|\\\verb|plural|$\to$ \verb|stem|$ + $ /\textipa{i}/};

%%     \node at (4.5,-5.5) {Polish Phonology};
%%     \node(polishPhonology) [draw,align = left] at (4.5,-6.3) {%
%%       $r_1$: \textipa{o}$\to$\textipa{u}$/$\verb|_[-nasal +voice]#|\\%
%%       $r_2$: \verb|[-sonorant]|$\to$\verb|[-voice]|$/$\verb|_#|};
%%     \node(polishMorphologyBox)[blue,draw,ultra thick, fit = (polishMorphologyLabel) (polishPhonology)] {};

%%     \draw[ultra thick,->] (universalBox.south) -- (ABAmorphologyBox.north);
%%     \draw[ultra thick,->] (universalBox.south) -- (polishMorphologyBox.north);

%%     \node(lexicon)[magenta,align = center,rotate = 90] at (1,-10) {\textbf{Words}};
%%     \draw[ultra thick,->] (morphology.west)-- (lexicon.east);

%%     \node(p1) [align = left] at (9.5,-8) {\verb|LATIN:| \textipa{l\ae tIn}};
%%     \node(p2) [draw,align = left] at (9.5,-9.5) {\textipa{l\ae tIn}\nextForm{r_1}\\\textipa{l\ae tInl}\nextForm{r_2}\\\textipa{l\ae tInle}\nextForm{r_3}\\\textipa{\ae tInle}}; %\\\textipa{l\ae tInle}}};
%%     \draw [thick,->] (p1.south) -- (p2.north);

%%     \node(w1) [align = left] at (12,-8) {\verb|ASK:| \textipa{\ae sk}};
%%     \node(w2) [draw,align = left] at (12,-9.5) {\textipa{\ae sk}\nextForm{r_1}\\\textipa{\ae sk}\nextForm{r_2}\\\textipa{\ae ske}\nextForm{r_3}\\\textipa{\ae ske}};
%%     \draw [thick,->] (w1.south) -- (w2.north);

%%     \node[magenta,draw,align= left](ABAvocabulary) at (10.5,-11.5){\textipa{\ae tInle}, \textipa{\ae ske}, $\cdots$ };
%%     \draw [thick,->] (p2.south) -- (ABAvocabulary.north);
%%     \draw [thick,->] (w2.south) -- (ABAvocabulary.north);
%%     \node(ABAlexiconBox)[draw,ultra thick, fit = (p1) (w1) (ABAvocabulary)] {};
%%     \draw [ultra thick,->] (ABAmorphologyBox.south) -- (ABAlexiconBox.north);

%%     \node(crib) at (4.5,-8) {\verb|horn:| \textipa{rog}};
%%     \node(Cingular)[draw,align = left] at (3.5,-9.5) {\verb|singular:|\\ \textipa{rog}\nextForm{r_1}\\\textipa{rug}\nextForm{r_2}\\\textipa{ruk}};
%%     \node(plural)[draw,align = left] at (5.5,-9.5) {\verb|plural:|\\ \textipa{rogi}\nextForm{r_1}\\\textipa{rogi}\nextForm{r_2}\\\textipa{rogi}};
%%     \draw [thick,->] (crib.south) -- (Cingular.north);
%%     \draw [thick,->] (crib.south) -- (plural.north);

%%     \node[magenta,draw,align = left] (polishObservation) at (4.5,-11.5) {\verb|horn, singular:| \textipa{\v{z}wop}\\\verb|horn, plural:| \textipa{\v{z}wubi}};
%%     \draw [thick,->] (Cingular.south) -- (polishObservation.north);
%%     \draw [thick,->] (plural.south) -- (polishObservation.north);

%%     \node(polishLexicon)[draw,ultra thick, fit = (crib) (Cingular) (plural) (polishObservation)] {};
%%     \draw [ultra thick,->] (polishMorphologyBox.south) -- (polishLexicon.north);

%%   \end{tikzpicture}
%% \end{figure}

%\bibliographystyle{named}
\bibliographystyle{unsrt} \bibliography{main}

>>>>

\usepackage{amsmath}
\usepackage[margin=0.1in]{geometry}
\usepackage{booktabs}
\usepackage{tipa}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{tikz}
\usetikzlibrary{fit}
%\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{phonrule} 
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

@@

A goal of the grammar learner, whether it is a child, linguist, or computer program, is to
infer the grammar that gave rise to
a collection of utterances.
One way of
modeling this inference problem is to (1) place a prior distribution over grammars, for example, a distribution that puts more probability mass on shorter or simpler grammars; (2) equip each grammar with a
\emph{likelihood model} that specifies exactly how likely a grammar is to produce a given utterance;
and then (3) use Bayes's rule to work backwards from the
utterances to the grammar that was likely to produce them:
\begin{equation}
  \probability\left[\text{grammar}|\text{utterances} \right]\propto\underbrace{\probability\left[\text{utterances}|\text{grammar} \right]}_{\text{Likelihood}}\times\underbrace{\probability\left[\text{grammar} \right]}_{\text{Prior}}\label{posterior}
\end{equation}
We can think of the above equation as a recipe for
scoring how well a grammar explains a
collection of utterances.
Once we have a prior and a likelihood,
we know how to evaluate how well a grammar explains the data, modulo the assumptions implicit in the
prior and likelihood.
In this work we will use this probabilistic framing
to explore spaces of grammars
that an AGL learner could use to explain
the pronunciations of words.
Here the grammars will be SPE-style rewrite rules~\cite{chomsky1968sound},
commonly used in generative phonology,
but at a high level this framing can apply to
learning other aspects of grammar like semantics~\cite{piantadosi2011learning}, syntax~\cite{perfors2011learnability}, and morphology~\cite{tim}.

\noindent\textbf{SPE-style rewrite notation:} Each SPE-style rewrite rule is a function that both inputs and outputs a sequence of phonemes.
These rewrites are written as ``\phonb{input}{output}{left}{right}'' %$\text{input}\to\text{output }/\text{ left } \_ \text{ right}$'',
which means that ``input'' gets rewritten to ``output'' whenever ``left'' is to the left and ``right'' is to the right.
Some special symbols used in these rules are \# to mean the start or end of the input;
C/V to mean a consonant/vowel; $\sigma $ to mean a syllable;
and $\varnothing$ to mean the empty string.
For example, the rule
\phonb{$\sigma $}{$\varnothing$}{\#}{C}
deletes a syllable at the start of a word if it is followed by a consonant.
Rules can refer to sets of phonemes by writing down vectors of phonological features -- for example,
the rule \phon{[-son]}{[+voice]} rewrites segments that are -sonorant (e.g., \textipa{k,d,v,S,...}) to their voiced counterparts (e.g., \textipa{g,d,v,Z,...}).
Rules can also bind variables to phonemes or syllables,
which are written using subscripts;
for example,
\phonr{C$_i$}{$\varnothing$}{C$_i$}
%$\text{\texttt{C}}_i\to\varnothing / \_ \text{\texttt{C}}_i$
deletes a consonant if it is followed by the same consonant,
and
\phonb{$ \varnothing $}{$\sigma _i$}{#}{$\sigma _i$}
%$\varnothing\to\sigma_i / \text{\texttt{\#}}\_\sigma_i$
inserts a copy of a word initial syllable.
The special subscript ``0" means zero or more occurrences of the subscripted segment,
so for example \phonl{V}{[+hi]}{[+hi][ ]$_0$} means that vowels become +hi whenever there exists a +hi phoneme anywhere to the left.

\noindent \textbf{Prior \& Likelihood:} A simple and intuitive prior over grammars is one which penalizes longer grammars and favors parsimonious grammars, for example defining $\probability\left[\text{grammar} \right]\propto\exp\left(- \text{\# symbols in grammar} \right)$. An example of a likelihood model,
and the one we use here,
is just to count the number of extra bits or symbols needed to encode an utterance given the grammar\footnote{This is equivalent to a minimum description length (MDL) interpretation}.

@@

The fundamental hypothesis underlying AGL research is that
artificial grammar learning must engage some shared resource with first language acquisition --
and so we can gain insights into first language acquisition by
instead studying AGL in controlled experiments.
Our demonstration here is that
a BPL model can explain how a learner can quickly (e.g., from very little data) infer a grammar represented as SPE--style rules.
We collected grammars from AGL datasets~\cite{gerken2010infants,marcus1999rule,frank2011three}
and presented our BPL learner with data drawn from these grammars (Table~\ref{artificialGrammarTable}),
finding that our model automatically discovers the grammars that
 underly these data sets.
\begin{table}[h!]
\centering\begin{tabular}{lcll}
  \toprule
  Grammar&\begin{tabular}{c}
    Example input \\to learner
    \end{tabular}&Inferred grammar&Natural language analogues\\\midrule
  \begin{tabular}{l}
    ABA \\
  Marcus et al. 1999.     
    \end{tabular}&
  \begin{tabular}{l}
    wofewo\\
    lovilo\\
    fimufi
  \end{tabular}&
  \begin{tabular}{l}
    \phonb{$\varnothing $}{$\sigma _i$}{\#}{$\sigma \sigma _i$}
%$\varnothing\to \sigma_i/$\texttt{\#\_}$\sigma \sigma_i$
  \end{tabular}& Reduplication (eg, Tagalog)
  \\\midrule
  \begin{tabular}{l}
      ABB\\
  Marcus et al. 1999. 
    \end{tabular}&
  \begin{tabular}{l}
        wofefe\\
    lovivi\\
    fimumu
  \end{tabular}&
  \begin{tabular}{l}
    \phonb{$\varnothing $}{$\sigma _i$}{$\sigma _i$}{\#}
    %$\varnothing\to \sigma_i/\sigma_i$\texttt{\_\#}
  \end{tabular}&
  Reduplication (eg, Tagalog)\\\midrule
  \begin{tabular}{l}
      ABx 
    \end{tabular}&
  \begin{tabular}{l}
        wofeka\\
    lovika\\
    fimuka
  \end{tabular}&
  \begin{tabular}{l}
stem$+x$
  \end{tabular}&concatenative morphology
  \\\midrule
  \begin{tabular}{l}
    AAx \\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wowoka\\
    loloka\\
    fifika
  \end{tabular}&
  \begin{tabular}{l}
    stem+$x$\\
    \phonb{$\varnothing $}{$\sigma _i$}{\#}{$\sigma _i$}
%$\varnothing\to \sigma_i/$\texttt{\#\_}$\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    reduplication\\concatenative morphology
    \end{tabular}\\\midrule
  \begin{tabular}{l}
    AxA \\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wokawo\\
    lokalo\\
    fikafi
  \end{tabular}&
  \begin{tabular}{l}
    $x+$stem\\
    \phonb{$\varnothing $}{$\sigma _i$}{\#}{$\sigma \sigma _i$}
    %$\varnothing$ $\to\sigma_i/$\texttt{\#\_}  $\sigma$ $\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    Infixing\\Reduplication
    \end{tabular}
  \\\midrule
  \begin{tabular}{l}
    Pig Latin
    \end{tabular}&
  \begin{tabular}{l}
    \textipa{pIg}$\to$\textipa{Igpe}\\
    \textipa{latIn}$\to$\textipa{atIle}\\
    \textipa{\ae sk}$\to$\textipa{\ae ske}\\
  \end{tabular}&
  \begin{tabular}{l}
    \phonb{$\varnothing $}{C$_i$}{\#C$_i$[ ]$_0$}{\#}\\
    \phonr{$\varnothing $}{\textipa{e}}{\#}\\
    \phonl{C}{$\varnothing $}{\#}
 %%      $\varnothing\to$\texttt{C}$_i$ $/$\texttt{\#C}$_i$\texttt{[ ]*\_\#}\\%
 %%      $\varnothing\to$ \textipa{e} $/$\texttt{\_\#}\\%
 %% \texttt{C}$\to\varnothing$ $/$ \texttt{\#\_}
  \end{tabular}  & \begin{tabular}{l}
    \end{tabular}
\\  \bottomrule  \end{tabular}
\caption{Using BPL to model artificial grammar learning. Learner given 5 examples.}\label{artificialGrammarTable}
\end{table}

@@

As an example of how the Pareto front offers a lens on the
space of possible hypotheses,
consider the upper left Pareto front in figure~\ref{quadrupleFrontier} (\emph{aab, 1 example}).
Here the learner has seen a single word, [\textipa{vEfefe}].
Some grammars living on the Pareto frontier are:
\begin{itemize}
\item \textbf{A grammar that does nothing:} In the lower right-hand corner of~\ref{quadrupleFrontier}
  is the grammar labeled ``surface=underlying'',
  which just says that every word (a ``surface pronunciation'')
  is represented (``underlyingly'') literally how it is pronounced.
  The observed word [\textipa{vEfefe}] is represented as /\textipa{vEfefe}/, which has 6 symbols,
  giving a fit to the data of -6.
\item \textbf{A grammar that duplicates syllables:} Highlighted in dark red in~\ref{quadrupleFrontier}
  is a grammar with the single rule
  \phonr{$\varnothing $}{  $\sigma _i$}{$\sigma _i$\#}
  %$\varnothing\to\sigma_i/\_\sigma_i$\texttt{\#},
  which inserts a copy of the last syllable syllable.
  This rule generates the word [\textipa{vEfefe}] by starting with the stem (i.e., underlying form) /\textipa{vEfe}/ (which has 4 symbols, fit to data of -4) and then applying the rule in the grammar,
  which copies the last syllable and makes [\textipa{vEfefe}].
\item \textbf{A grammar that duplicates syllables and appends
  morphemes:} Highlighted in pink in~\ref{quadrupleFrontier} are
  grammars that duplicate a syllable but also append or prepend extra
  morphemes.  These correspond to generalizations where the learner
  believes that different parts of [\textipa{vEfefe}] correspond to
  prefixes or suffixes in the language.  For example, the grammar in
  the upper left corner incorporates the suffix /\textipa{fe}/ as well
  as the prefix /\textipa{vE}/, and explains the observation using an
  underlying form that is completely empty (fit to data of 0) -- this
  grammar has memorized the observed data, and so maximally compresses
  it, but at the cost of having many symbols inside of the grammar (22 symbols, vs 6 symbols in the grammar that just duplicates a syllable).  
  \end{itemize}

@@

\includegraphics[width = 0.5\textwidth]{alternativeHypothesesFigures/abb1.png} \includegraphics[width = 0.5\textwidth]{alternativeHypothesesFigures/abb3.png}
\includegraphics[width = 0.5\textwidth]{alternativeHypothesesFigures/aax1.png} \includegraphics[width = 0.5\textwidth]{alternativeHypothesesFigures/aax3.png}

@@

These natural language results are still preliminary and
ongoing. Figure~\ref{naturalLanguageExample} shows our model solving a
representative textbook problem.  We believe this extension to
natural language is an important next step for models of artificial
grammar learning, and helps deliver on the basic promises and
presuppositions of AGL research.

\begin{figure}[H]\centering
  \begin{minipage}{7cm}\centering
    \textbf{Textbook Tibetan counting problem:}
    \begin{tabular}{ll}\toprule
Number & Surface form
\\ \midrule
10 & \textipa{d\super Zu}\\
1 & \textipa{d\super Zig}\\
11 & \textipa{d\super Zugd\super Zig}\\
4 & \textipa{Si}\\
%14 & \textipa{d\super ZubSi}\\
40 & \textipa{Sibd\super Zu}\\
9 & \textipa{gu}\\
19 & \textipa{d\super Zurgu}\\
%90 & \textipa{gubd\super Zu}\\
5 & \textipa{Na}\\
%15 & \textipa{d\super ZuNa}\\
%50 & \textipa{Nabd\super Zu}\\
\bottomrule
\end{tabular}
  \end{minipage}%
  \hspace{0.25cm}
  \begin{minipage}{7cm}\centering
    \textbf{Our model's solution:}

    Induced phonological rule: \phonb{C}{$\varnothing $}{\#}{C} \\ %\texttt{C}$\to\varnothing /$\texttt{\#\_C} \\
    \hspace{0.5cm}\emph{(Reduce word initial consonant clusters)}

@@

\bibliographystyle{unsrt} \bibliography{main}
